---
title: "Assignment2"
author: "Jing Xu"
date: "2022-10-05"
output: html_document
---

## Load Package

```{r}
library(tidyverse)
# install.packages("Stat2Data")
library(Stat2Data)
Hawks <- data("Hawks")
```

# Data Wrangling

## 1.1 Select and filter

### (Q1)

```{r}
# head(Hawks)
hSF <- Hawks %>% filter(Species == "RT" & Weight >= 1000) %>% select(Wing, Weight, Tail)
head(hSF)
```

### (Q2)

```{r}
# how many variables in hSF
varNum <- ncol(hSF)
cat("Number of variables in hSF :", varNum)
rowNum <- nrow(hSF)
cat("Number of examples in hSF :", rowNum)
```

## 1.2 The arrange function

### (Q1) Use the arrange() function to sort the hSF data frame

```{r}
ascHSF <- hSF %>% arrange(Wing)
head(ascHSF)
```

## 1.3  Join and rename functions

### Q1

```{r}
species_code <- c("CH", "RT", "SS")
species_name_full <- c("Cooper's", "Red-tailed", "Sharp-shinned")
hawkSpeciesNameCodes <- data.frame(species_code, species_name_full)
```

### Q2

```{r}
# head(Hawks)
hawksFullName <- left_join(Hawks, hawkSpeciesNameCodes, by=c("Species" = "species_code")) %>% 
  select(-Species) %>% 
  rename(Species = species_name_full)
head(hawksFullName)
```

### Q3

```{r}
head(hawksFullName, n=7) %>% select(Species, Wing, Weight)
```

## 1.4  The mutate function

### Q1

```{r}
hawksWithBMI <- Hawks %>% 
  mutate(bird_BMI=1000*Weight/(Wing)^2) %>% 
  select(Species, bird_BMI) %>% 
  arrange(desc(bird_BMI))
head(hawksWithBMI, n=8)
?abs
```

## 1.5 Summarize and group-by functions

### Q1

```{r}
#?mean
#?median
hawksFullName %>%
  group_by(Species) %>%
  summarize(num_rows=n(), mn_wing=mean(Wing, na.rm=TRUE), md_wing=median(Wing, na.rm=TRUE), t_mn_wing=mean(Wing, trim=0.1), b_wt_ratio=max(Wing/Tail, na.rm=TRUE))
```

### Q2

```{r}
#?across
hawksFullName %>%
  select(Species, Wing, Weight, Culmen, Hallux, Tail, StandardTail, Tarsus, Crop) %>%
  group_by(Species) %>%
  summarize(across(everything(), ~sum(is.na(.x))))
```

# 2. Tidy data and iteration

## 2.1 Missing data and iteration

### Q1

We need to load 'tidyverse' library.

### Q2

```{r}
impute_by_median <- function(x) {
  md <- median(x, na.rm = 1)
  impute_f <- function(z) {
    if (is.na(z)) {
      return (md)
    } else {
      return (z)
    }
  }
  return (map_dbl(x, impute_f))
}
v<-c(1,2,NA,4)
impute_by_median(v)
```

### Q3

```{r}
?map
?list
x<-seq(0, 10, 0.1)
y<-seq(1, 51, 0.5)
df_xy <- data.frame(x,y)
map2sum <- map2_dbl(x, y, ~ .x+.y)
res <- data.frame(x,y,map2sum)
head(res)
```

### Q4

```{r}
sometimes_missing <- function(idx, val) {
  if (idx%%5 == 0) {
    return (NA)
  } else {
    return (val)
  }
}
sometimes_missing(14,25)
sometimes_missing(15,25)
#?data.frame
df_xy_missing <- data.frame(x,y) %>% mutate(y=map2_dbl(row_number(y), y, sometimes_missing))
head(df_xy_missing, n=10)
```

### Q5

```{r}
df_xy_imputed <- df_xy_missing %>% mutate(y=impute_by_median(df_xy_missing$y))
head(df_xy_imputed)
```

## 2.2 Tidying data with pivot functions

```{r}
#install.packages("readxl")
library(readxl)
library(tidyverse)
wins_data_frame <- read_excel("/Users/xujing/Documents/DS/SCEM/SCEM/HockeyLeague.xlsx")
wins_data_frame %>%
  select(1:5) %>%
  head(3)
```

### Q1

```{r}
wins_tidy <- wins_data_frame %>%
  rename(Team=1) %>%
  pivot_longer(2:ncol(wins_data_frame), names_to = 'Year', values_to = 'value') %>%
  separate(value, into = c('Wins', 'Total'), sep = 'of', convert = TRUE)
wins_tidy %>% dim()
head(wins_tidy, n=5)
```

### Q2

```{r}
losses_data_frame <- read_excel("/Users/xujing/Documents/DS/SCEM/SCEM/HockeyLeague.xlsx", "Losses")
losses_tidy <- losses_data_frame %>%
  rename(Team=1) %>%
  pivot_longer(2:ncol(losses_data_frame), names_to = 'Year', values_to = 'value') %>%
  separate(value, into = c('Losses', 'Total'), sep = 'of', convert = TRUE)
losses_tidy %>% dim()
head(losses_tidy, n=5)
```

### Q3

```{r}
hockey_df <- wins_tidy %>% 
    left_join(losses_tidy %>% select(-Total), by=c("Team" = "Team", "Year" = "Year")) %>% 
    mutate(Draw=Total-Wins-Losses, Wins_rt=Wins/Total, Losses_rt=Losses/Total, Draws_rt=Draw/Total)
hockey_df %>% head(5)
```

### Q4

```{r}
hockey_df %>%
    select(Team, Wins_rt, Losses_rt, Draws_rt) %>%
    rename(W = Wins_rt, L = Losses_rt, D = Draws_rt) %>%
    group_by(Team) %>%
    summarize(across(everything(), ~median(.), .names = "{.col}_md"), across(W:D, ~mean(.), .names = "{.col}_mn")) %>%
    arrange(desc(W_md))
```

# 3. Visualisation

## Q1

```{r}
library(ggplot2)
wins_plot = wins_tidy %>%
    filter(Team=="Ducks") %>%
    ggplot(aes(x=Wins)) + xlab("wins") 
wins_plot + geom_histogram(binwidth = 3) + ylab("Count")  
```

## Q2

The parameter 'adjust' in the function 'geom_density' control the smoothness of the curve, the bigger value of the adjust, the more smooth the curve is.

```{r}
wins_plot + geom_density(adjust=0.5) + ylab("Count")
wins_plot + geom_density(adjust=2) + ylab("Count")
```

## Q3

```{r}
?pivot_wider
wins_teams <- wins_tidy %>%
    select(-Total) %>%
    pivot_wider(names_from = Team, values_from = Wins)
head(wins_teams)
wins_teams %>% ggplot(aes(y=Eagles, x=Ducks)) + xlab("Ducks") + ylab("Eagles") + geom_point()
```






